A template for the *arxiv* style

### 2022-02-06 2-the-cubble-pkg.Rmd
all the  -> \code{CODE} as per jss  guideline

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data. A spatial identifier (`id` in the diagram) is the unique identifier of each site. The temporal identifier (`t` in the diagram) prescribes the time stamp each site is recorded. Coordinates, comprising of latitude and longitude (`lon` and `lat` in the diagram), locates each site on the map. These identifiers will be the building blocks for the data structure introduced below. Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e. the name or code of the weather station, and temporal variables that varies with time.

@@

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes. The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e. the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form). Underneath the nested form, a cubble is built from a row-wise dataframe (`rowwise_df`) where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of `map` syntax when working with list-column.

@@

`as_cubble()` creates a \pkg{cubble} by supplying the three key components: `key` as the spatial identifier; `index` as the temporal identifier; and a vector of `coords` in the order of longitude first and then latitude. The naming of `key` and `index` follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

@@

There are a few information in the \pkg{cubble} header: the name of the `key` variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the `ts` column with its type: `date [date], prcp [dbl], tmax [dbl], tmin [dbl]`.

@@

The long cubble is suitable to manipulate the time dimension of the data. The function `stretch()` switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the `spatial` attribute and then unnests the `ts` column:

@@

Notice here that the third line in the header now shows the name and type of spatial variables: `lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]`.

@@

Manipulation on the spatial and temporal dimension can be an iterative process. Many times, analysts will need to go back and forth between the nested and long cubble. `tamp()` inverses the `stretch()` that introduced in the previous section, to switch from a long cubble to the nested cubble:

@@

Some functions require time invariant variables to be in the same table as the time varying ones. In cubble, this can be done through `migrate()`, which moves the time invariant variables into the long cubble:

@@

Building from an `tbl_df` class, `cubble` has implemented methods for `dplyr` generics, which includes:

@@

\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the `coords` argument need to be specified:

@@

In the created nested form, each element of the list-column `ts` is of `tbl_ts` class so that existing operations on the tsibble class can still be applied:

>>>>

Spatio-temporal data usually come in various forms and Figure \ref{fig:illu-input} shows four examples of this. No matter which form the data is in, these formats share some common components that characterise the spatio-temporal data. A spatial identifier (\code{id} in the diagram) is the unique identifier of each site. The temporal identifier (\code{t} in the diagram) prescribes the time stamp each site is recorded. Coordinates, comprising of latitude and longitude (\code{lat} and \code{(long)} in the diagram), locates each site on the map. These identifiers will be the building blocks for the data structure introduced below. Other variables in the data can be categorised into two groups: spatial variables that are invariant at each time stamp for every site, i.e. the name or code of the weather station, and temporal variables that varies with time.

@@

In a cubble, there are two forms: nested form and long form, and Figure \ref{fig:illu-cubble} sketches the two forms along with the associated attributes. The decision on which form to use is output-oriented, meaning analysts need to first think about whether the output of a particular operation is identified only by the spatial identifier, or a combination of spatial and temporal identifier. The nested cubble is suitable for working with operations that are only identified by the site and this type of operation can be a pure manipulation of spatial variables, or a summary of temporal variables by site (i.e. the output of counting the number of raining day is only identified by sites and hence should be performed with the nested form). Underneath the nested form, a cubble is built from a row-wise dataframe (\code{rowwise_df}) where each site forms a separate group. This structure simplifies the calculation that involves temporal variables by avoiding the use of \code{map} syntax when working with list-column.

@@

`as_cubble()} creates a \pkg{cubble} by supplying the three key components: \code{key} as the spatial identifier; \code{index} as the temporal identifier; and a vector of \code{coords} in the order of longitude first and then latitude. The naming of \code{key} and \code{index} follows the convention in the \pkg{tsibble} package. The cubble created by default is in the nested form. Below is an example of creating a cubble:

@@

There are a few information in the \pkg{cubble} header: the name of the \code{key} variable (`id`) and its number (5), bounding box (`[115.97, -32.94, 133.55, -12.42]`), and also the name of variable nested in the \code{ts} column with its type: \code{date [date], prcp [dbl], tmax [dbl], tmin [dbl]}.

@@

The long cubble is suitable to manipulate the time dimension of the data. The function \code{stretch()} switches the nested cubble into the long cubble by first extracts all the spatial variables into a separate tibble and store in the \code{spatial} attribute and then unnests the \code{ts} column:

@@

Notice here that the third line in the header now shows the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}.

@@

Manipulation on the spatial and temporal dimension can be an iterative process. Many times, analysts will need to go back and forth between the nested and long cubble. \code{tamp()} inverses the \code{stretch()} that introduced in the previous section, to switch from a long cubble to the nested cubble:

@@

Some functions require time invariant variables to be in the same table as the time varying ones. In cubble, this can be done through \code{migrate()}, which moves the time invariant variables into the long cubble:

@@

Building from an \code{tbl_df} class, \pkg{cubble} has implemented methods for \pkg{dplyr} generics, which includes:

@@

\pkg{cubble} is also compatible with \pkg{tsibble}. When creating a cubble from a tsibble, Only the \code{coords} argument need to be specified:

@@

In the created nested form, each element of the list-column \code{ts} is of \code{tbl_ts} class so that existing operations on the tsibble class can still be applied:

### 2022-02-06 3-advanced.Rmd
all the  -> \code{CODE} as per jss  guideline

Spatial locations can have grouping structure that is either inherent to the data  (i.e. countries are nested in continents), or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, `switch_key()` can be used to create a new level of grouping of spatial location by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying `cluster_nested <- station_nested %>% switch_key(key = cluster)`, the cubble re-defines the cubble key from `id` in `station_nested` to `cluster` in `cluster_nested`. All the spatial variables variant to `cluster` are now nested into a `.val` column and cluster level variables can be computed in the same fashion as station level variables in `station_nested`. Following the same principal of `stretch`, the long form cluster level cubble expands the time series variables with both indices: `cluster` and `id`. All the cluster level variables are stored separately for recovering the nested from from the long form.

@@

`match_sites()` from `cubble` matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series `A` and `a`, 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series `A` and the peaks in series `a` are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in `match_sites()` to fine-tune the matching:

@@

 - `spatial_n_keep`: the number of spatial match for each site to keep
 - `spatial_dist_max`: the maximum distance allowed for a matched pair 
 - `temporal_n_highest`: the number of peak used - 3 in the example above
 - `temporal_window`: the length of the interval - 5 in the example above
 - `temporal_min_match`: the minimum number of matched peak for a valid matched pair

@@

Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series with the two forms of a cubble. The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left). The nested cubble will communicate to the long cubble to activate all the observations with the same `id` (middle). The long cubble will then highlight the activated series in the time series plot (right).

The linking is also available from the time series plot to the map. The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble. All the observations that share the same `id` are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble. These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, `GGally` implements the glyph map through the `glyphs()` function. The function construct a data frame with calculated position (`gx`, `gy`, `gid`) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into `ggplot` to create the glyph map as:

@@

A re-implementation of the glyph map as a ggproto `GeomGlyph` has been made in the `cubble` package so that the glyph map can be created with `geom_glyph()`:

@@

Polar glyph map can be specify as a parameter, `polar = TRUE`, in the `geom_glyph()`, along with `width` and `height` in either absolute or relative value. Global and local scale can be controlled by the parameter `global_rescale`, which default to `TRUE` for global scaling. Reference box and line can be added with separate `geom_glyph_box()` and `geom_glyph_line()`.

>>>>

Spatial locations can have grouping structure that is either inherent to the data  (i.e. countries are nested in continents), or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial location by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}. All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}. Following the same principal of \code{stretch}. the long form cluster level cubble expands the time series variables with both indices: \code{cluster} and \code{id}. All the cluster level variables are stored separately for recovering the nested from from the long form.

@@

`match_sites()} from \pkg{cubble} matches data based on these two criteria. It first matches the two data sources spatially through computing the pairwise distance based on latitude and longitude. Pairs that pass the spatial matching are then matched temporally through computing the number of matched peaks within a fixed length window. Figure \ref{fig:illu-matching} illustrate this temporal matching in more details. Given two series \code{A} and \code{a}. 3 peaks have been picked in each series. An interval, with default length of 5, is constructed for each peak in series \code{A} and the peaks in series \code{a} are tested against whether they fall into the any of the intervals. In this illustration, there are 2 matches for these two series. Several arguments are available in \code{match_sites()} to fine-tune the matching:

@@

 - \code{spatial_n_keep}. the number of spatial match for each site to keep
 - \code{spatial_dist_max}. the maximum distance allowed for a matched pair 
 - \code{temporal_n_highest}. the number of peak used - 3 in the example above
 - \code{temporal_window}. the length of the interval - 5 in the example above
 - \code{temporal_min_match}. the minimum number of matched peak for a valid matched pair

@@

Cubble fits in naturally with the interactive graphic pipeline discussed in the literature [@buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling]. Diagram \ref{fig:illu-interactive} illustrates how linking works from the map to the time series with the two forms of a cubble. The map and time series plot is associated with the nested or long cubble respectively and when a user action is captured on the map, the site will be activated in the nested cubble (left). The nested cubble will communicate to the long cubble to activate all the observations with the same \code{id} (middle). The long cubble will then highlight the activated series in the time series plot (right).

The linking is also available from the time series plot to the map. The selection(s) on the time series is through selecting the point(s) on the time series and once a point is selected, it will be activated in the long cubble. All the observations that share the same \code{id} are then activated and this includes other points in the same time series in the long cubble and the corresponding observation of site in the nested cubble. These activated observations will then being reflected in the updated plots and Diagram \ref{fig:illu-interactive-2} in the Appendix illustrates this process.

@@

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function construct a data frame with calculated position (`gx}. \code{gy}. \code{gid}. of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

A re-implementation of the glyph map as a ggproto \code{GeomGlyph} has been made in the \pkg{cubble} package so that the glyph map can be created with \code{geom_glyph()}:

@@

Polar glyph map can be specify as a parameter, \code{polar = TRUE}. in the \code{geom_glyph()}. along with \code{width} and \code{height} in either absolute or relative value. Global and local scale can be controlled by the parameter \code{global_rescale}. which default to \code{TRUE} for global scaling. Reference box and line can be added with separate \code{geom_glyph_box()} and \code{geom_glyph_line()`.

### 2022-02-06 4-1-basic.Rmd
all the  -> \code{CODE} as per jss  guideline

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset `weatherdata::historical_tmax` extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations. All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`. The data `historical_tmax` is already presented as a cubble, with `id` as the key, `date` as the index, and `c(longitude, latitude)` as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

@@

Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by `stretch()`:

@@

There are a few stations which don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using `tamp()`:

@@

Lastly, to create a glyph map, both the major (`longitude`, `latitude`) and minor (`month`, `tmax`) coordinates need to be in the same table. Spatial variables can be moved to the long form with `migrate()`:

@@

`tmax` can then be supplied to `geom_glyph()` for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:

>>>>

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world and the dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations. All the stations start recording from year `r range(weatherdata::historical_tmax$first_year)[2]` with the longest back to year `r range(weatherdata::historical_tmax$first_year)[1]`. The data \code{historical_tmax} is already presented as a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

@@

Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:

@@

There are a few stations which don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:

@@

Lastly, to create a glyph map, both the major (\code{longitude}, \code{latitude}) and minor (\code{month}, \code{tmax}) coordinates need to be in the same table. Spatial variables can be moved to the long form with \code{migrate()}:

@@

\code{tmax} can then be supplied to \code{geom_glyph()} for the glyph map in Figure \ref{fig:basic-manip} with a station inset on the top left corner:

### 2022-02-06 4-2-agg.Rmd
all the  -> \code{CODE} as per jss  guideline

This example shows how to organise data at both level with `switch_key()`. The dataset used is `weatherdata::climate_full`, which records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used here to create 20 clusters. This creates `station_nested` as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

To create a group level cubble, use `switch_key()` with the new key variable, `cluster`:

@@

With the group level cubble, `get_centroid()` is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

>>>>

This example shows how to organise data at both level with \code{switch_key()}. The dataset used is \code{weatherdata::climate_full}, which records daily precipitation and maximum/ minimum temperature for \code{r nrow(weatherdata::climate_full)} stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used here to create 20 clusters. This creates \code{station_nested} as a station level nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problem, as long as there is a mapping from each station to a cluster.

@@

To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:

@@

With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

### 2022-02-06 4-3-water-level.Rmd
all the  -> \code{CODE} as per jss  guideline

From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station. As introduced in Section 3.2, `match_sites()` can be used to match one source of data with another source in a cubble. Here `prcp` in `climate` will be matched to `Water_course_level` in `river`, which is specified in the argument `temporal_var_to_match`. `temporal_independent` controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this). Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation. This will put precipitation data, `climate`, as the independent. Given there is one year worth of data, the number of peak (`temporal_n_highest`) to consider is slightly raised from a default 20 to 30. `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`

@@

The output from matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the temporal matching.

>>>>

From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{prcp} in \code{climate} will be matched to \code{Water_course_level} in \code{river}, which is specified in the argument \code{temporal_var_to_match}. \code{temporal_independent} controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this). Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation. This will put precipitation data, \code{climate}, as the independent. Given there is one year worth of data, the number of peak (\code{temporal_n_highest}) to consider is slightly raised from a default 20 to 30. \code{temporal_min_match} filters out pairs don't have enough match and to return all the pairs, set \code{temporal_min_match} to 0.

@@

The output from matching is also a cubble, with additional column \code{dist} and \code{group} produced from spatial matching and \code{n_match} from the temporal matching.

### 2022-02-06 4-4-netcdf.Rmd
all the  -> \code{CODE} as per jss  guideline

NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: dimension to define the spatiotemporal grid (longitude, latitude, and time) and variable which populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format. Current packages on manipulating NetCDF data in R include a high-level R interface: `ncdf4`[@ncdf4], a low-level interface that calls C-interface: `RNetCDF`[@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: `tidync`[@tidync].

Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`. It maps each combination of longitude and latitude into an `id` (the `key`) and nests time and variables in the nested form:

@@

Memory limit with Netcdf data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution. Subsetting longitude and latitude grid is available through `long_range` and `lat_range` if the Netcdf file has finer resolution than needed.

```{r eval = FALSE, echo = TRUE}
# Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
# at 0.25 degree resolution and subset it to have 
# 1 degree resolution:
dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format. It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr]. Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded, the data can be read into a cubble using the code above. Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.

>>>>

NetCDF (Network Common Data Form) is a data format commonly used in climatology community to deliver global mapping of atmosphere, ocean, and land. It has two major components: dimension to define the spatiotemporal grid (longitude, latitude, and time) and variable which populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise this data format. Current packages on manipulating NetCDF data in R include a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

Cubble provides an \code{as_cubble()} method to coerce the \code{ncdf4} class from the \pkg{ncdf4} package into a \code{cubble}. It maps each combination of longitude and latitude into an \code{id} (the \code{key}) and nests time and variables in the nested form:

@@

Memory limit with Netcdf data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution. Subsetting longitude and latitude grid is available through \code{long_range} and \code{lat_range} if the Netcdf file has finer resolution than needed.

```{r eval = FALSE, echo = TRUE}
# Assume \code{my_ncdf} has bounding box of [-180, -90, 180, -90] 
# at 0.25 degree resolution and subset it to have 
# 1 degree resolution:
dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format. It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package \pkg{ecmwfr} [@ecwmfr]. Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 and once downloaded, the data can be read into a cubble using the code above. Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.

### 2022-02-06 4-5-interactive.Rmd
all the  -> \code{CODE} as per jss  guideline

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series. Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with `cubble` and `crosstalk`.

Starting with the original data, some pre-processing may be required to summarise the data before the visualisation. This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between `tmax` and `tmin`. The three daily variables are then averaged into month over 2016 - 2020 in the long form. Variance of the temperature difference is then calculated for each station in the nested form. Then, a `SharedData` object is constructed for each form of the cubble and the same `group` argument ensures the cross-linking of the two forms via the common `id` column. The spatial map and time series plot are then made with each `SharedData` objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. And the cross-linking is always safeguarded by the shared `id` column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map. This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with `leafpop::addPopupGraphs()`:

>>>>

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series. Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk{.

Starting with the original data, some pre-processing may be required to summarise the data before the visualisation. This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between \code{tmax} and \code{tmin}. The three daily variables are then averaged into month over 2016 - 2020 in the long form. Variance of the temperature difference is then calculated for each station in the nested form. Then, a \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. And the cross-linking is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map. This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

### 2022-02-06 main.Rmd
fix marked red sentences labelled for myseyself
all the  -> \code{CODE} as per jss  guideline

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    department: Department of Econometrics and Business Statistics
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolas@langrene.com
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Sptio-temporal data naturally has two levels of observations: site and time level. [problem]. In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
  plain:     [spatio temporal, R, climate weather station, data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: rticles::jss_article
bibliography: references.bib
---

@@

```
\newpage

>>>>

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolas@langrene.com
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Sptio-temporal data naturally has two levels of observations: site and time level. [problem]. In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
  plain:     [spatio temporal, R, climate weather station, data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: rticles::jss_article
bibliography: references.bib
---

@@
